Learning and comparing functional connectomes across subjects

### 2013-01-02 paper.tex
iter

% General introduction
Functional connectivity captures the synchronization of distant
neural systems via correlations in brain activity
\cite{biswal1995,cordes2000}. Given that high-level
function emerges from the interaction of specialized units, functional
connectivity is an essential part of the description of brain function,
that complements the localizationist picture emerging from the systematic
mapping of regions recruited in tasks. However, while there exists a
well-defined standard analysis framework for activation mapping that enables
statistically-controlled comparisons across subjects \cite{friston1995},
group-level analysis of functional connectivity still faces many open
methodological challenges. Giving a picture of a single subject's 
functional connectivity is by itself not straightforward, as the brain
comprises a myriad of interacting subsystems and its connectivity must 
be decomposed in
simplified and synthetic representations. An important view of brain
connectivity is that of distributed functional networks depicted by their
spatial maps \cite{fox2005}. Another no less important and complementary
view is that of connections linking localized functional modules depicted
as a graph \cite{bullmore2009}. This representation of brain connectivity
is often called the functional connectome \cite{sporns2005} and is the
focus of intense worldwide research efforts as it holds promises of new
insights in cognition and pathologies
\cite{greicius2008b,biswal2010,fox2010}.

@@

Horwitz \emph{el al.}\ \cite{horwitz1995} claimed almost 20 years ago
that ``the crucial concept needed for network analysis is covariance''.
In our eyes, this still holds today. Estimation functional connectomes
relies largely on fitting covariance models. Their comparison requires
understanding how these covariances vary and finding metrics to capture
this variability. The additional secret ingredient may be a good choice
of confounds in all statistical steps. A good choice of a small number of
relevant regions facilitates connectome comparison. However, such a
choice cannot yet be fully factored out via methods and must rely on
neuroscientific expertise.

>>>>

% General introduction
Functional connectivity reveals the synchronization of distant
neural systems via correlations in brain activity
\cite{biswal1995,cordes2000}. Given that high-level
function emerges from the interaction of specialized units, functional
connectivity is an essential part of the description of brain function,
that complements the localizationist picture emerging from the systematic
mapping of regions recruited in tasks. However, while there exists a
well-defined standard analysis framework for activation mapping that enables
statistically-controlled comparisons across subjects \cite{friston1995},
group-level analysis of functional connectivity still faces many open
methodological challenges. Giving a picture of a single subject's 
functional connectivity is by itself not straightforward, as the brain
comprises a myriad of interacting subsystems and its connectivity must 
be decomposed in
simplified and synthetic representations. An important view of brain
connectivity is that of distributed functional networks depicted by their
spatial maps \cite{fox2005}. Another no less important and complementary
view is that of connections linking localized functional modules depicted
as a graph \cite{bullmore2009}. This representation of brain connectivity
is often called the functional connectome \cite{sporns2005} and is the
focus of intense worldwide research efforts as it holds promises of new
insights in cognition and pathologies
\cite{greicius2008b,biswal2010,fox2010}.

@@

Horwitz \emph{el al.}\ \cite{horwitz1995} claimed almost 20 years ago
that ``the crucial concept needed for network analysis is covariance''.
In our eyes, this still holds today. Estimation functional connectomes
relies largely on fitting covariance models. Their comparison requires
understanding how these covariances vary and finding metrics to capture
this variability. The additional secret ingredient may be using  
confounds regressors in all statistical steps. A good choice of a small number of
relevant regions facilitates connectome comparison. However, such a
choice cannot yet be fully factored out via methods and must rely on
neuroscientific expertise.
